---
title: "Final Project/-608"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Headers

Import libraries to manipulate and clean the dataset

```{r warning=FALSE, message=FALSE}
library(tidyr)
library(dplyr)
library(ggplot2)
library(sqldf)
library(reshape)
library(zipcode)
```

## Get Data 

Load data to the R environment 

```{r}
EffectiveCaredataHospital <- read.csv(file="/Users/Admin/Documents/R/Timely and Effective Care Hospital.csv", header=TRUE, sep=",")
```

## Clean Data

Drop none needed column fields

```{r}
EffectiveCaredata <- select(EffectiveCaredataHospital,-Measure.Start.Date,-Measure.End.Date,-Footnote,-Measure.ID)
```

Find the measure variables

```{r}
EffectiveCaredata$Measure.Name[!duplicated(EffectiveCaredata$Measure.Name)]
```

Select the needed Variable for measure regarding waiting time 

Door to diagnostic eval: waiting for diagnostic 
ED1: diagnostic time of admitted patients 
Emergency department volume: Number of patients per year
Ed2: admission waiting time for getting a room 
Left before being seen: percentage of patient did not get diagnostic
opt18: diagnostic time of discharged patients

```{r warning=FALSE, message=FALSE}

EffectiveCaredata<-filter(EffectiveCaredata, Measure.Name == "ED1" |
         Measure.Name == "ED2" |
         Measure.Name == "Emergency department volume" |
         Measure.Name == "Door to diagnostic eval" |
         Measure.Name == "Left before being seen" |
         Measure.Name == "OP 18")
co=c(EffectiveCaredata$Hospital.Name,EffectiveCaredata$Provider.ID)

EffectiveCaredata=select(EffectiveCaredata,Measure.Name,Sample,Provider.ID,Hospital.Name,Address,City,State,ZIP.Code,County.Name,Score)

```

Change the categorical volume into numeric level  

```{r}
EffectiveCaredata$Score[EffectiveCaredata["Score"] == "Very High (60,000+ patients annually)"]<-1
EffectiveCaredata$Score[EffectiveCaredata["Score"] == "High (40,000 - 59,999 patients annually)"]<-2
EffectiveCaredata$Score[EffectiveCaredata["Score"] == "Medium (20,000 - 39,999 patients annually)"]<-3
EffectiveCaredata$Score[EffectiveCaredata["Score"] == "Low (0 - 19,999 patients annually)"]<-4

```

Pivot the measure variables into columns and order by zipcode
```{r warning=FALSE, message=FALSE}
EffectiveCaredata<-cast(EffectiveCaredata, Provider.ID+Hospital.Name+State+City+County.Name+ZIP.Code+Address~Measure.Name)
EffectiveCaredata=as.data.frame(EffectiveCaredata)
EffectiveCaredata<-EffectiveCaredata[order(EffectiveCaredata$ZIP.Code),]
row.names(EffectiveCaredata) <- NULL
```

Load the zip code dataset to get latitude and longitude of hospitals locations.

```{r}
data(zipcode)
colnames(zipcode)[1]<-"ZIP.Code"
zipcode<-transform(zipcode, ZIP.Code = as.numeric(ZIP.Code))
zipcode<-transform(zipcode, city = toupper(city))
```

Get the latitude and lontitude for all hospitals

```{r}

tempdatframe<-data.frame()

for(i in 1:nrow(EffectiveCaredata)){
  temp<-zipcode[zipcode$ZIP.Code==EffectiveCaredata[i,6],]
  tempdatframe<-rbind(tempdatframe,temp)
}

tempdatframe<-select(tempdatframe,-ZIP.Code,-city,-state)

EffectiveCareWithLocation<-data.frame()
EffectiveCareWithLocation<-cbind(EffectiveCaredata,tempdatframe)
row.names(EffectiveCareWithLocation) <- NULL
colnames(EffectiveCareWithLocation)<-c("Provider.ID","Hospital","State","City","County","Zip","Address",
                                       "Door to diagnostic","Diagnostic time of admitted patients",
                                       "Admission waiting time for getting a room ","ER Volume",
                                       "Left before being seen","Diagnostic time of discharged patients",
                                       "Latitude","Longitude")

```

Save data to file EffectiveCareWaitingTime.csv
```{r}
write.csv(EffectiveCareWithLocation, file = "EffectiveCareWaitingTime.csv")
```
